---
title: "Case studies with CompAREdesign: Binary endpoint"
author: "Marta Bofill Roig, Jordi Cortés Martínez and Guadalupe Gómez Melis"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE, echo=FALSE}
library('CompAREdesign')            # Load library
knitr::opts_chunk$set(echo = TRUE)  # Show code
```

## Binary endpoint: TUXEDO TRIAL

This example is based on the data from the [TUXEDO trial](https://doi.org/10.1056/nejmoa1510188) [2].

### Input Parameters

First of all, the information for the components of the composite endpoint should be defined.

```{r}
## Probabilities of observing the event in control arm at the end of follow-up
p0_e1 <- 0.059    # Ischemia-driven target-lesion revascularization 
p0_e2 <- 0.032    # Cardiac death or target-vessel MI 

## Effect size (absolute reduction) for each endpoint
AR_e1 <- -0.0196  # Ischemia-driven target-lesion revascularization 
AR_e2 <- -0.0098  # Cardiac death or target-vessel MI

## Correlation
rho   <- 0.4
```

### ARE: Asymptotic Relatively Efficiency

As the ARE is greater than 1, we can state that the design using the composite endpoint is more efficient.

```{r}
ARE_cbe(p0_e1   = p0_e1  , p0_e2   = p0_e2, 
        eff_e1  = AR_e1  , eff_e2  = AR_e2, 
        effm_e1 = "diff" , effm_e2 = "diff", effm_ce = "or",
        rho     = rho) 
```

### Effect size of the Composite Endpoint

We can obtain the expected treatment effect based on the odds ratio (*OR*).

```{r}
effectsize_cbe(p0_e1   = p0_e1  , p0_e2   = p0_e2, 
               eff_e1  = AR_e1  , eff_e2  = AR_e2, 
               effm_e1 = "diff" , effm_e2 = "diff", effm_ce = "or",
               rho     = rho) 
```

### Sample size

The required sample size for the design usign the composite endpoint is **2644**.

```{r}
samplesize_cbe(p0_e1   = p0_e1  , p0_e2   = p0_e2, 
               eff_e1  = AR_e1  , eff_e2  = AR_e2, 
               effm_e1 = "diff" , effm_e2 = "diff", effm_ce = "or",
               rho     = rho,
               alpha   = 0.05, beta = 0.2) 
```

## References

1. Herbst RS, Sun Y, Eberhardt WEE, Germonpré P, Saijo N, Zhou C et al. Vandetanib plus docetaxel versus docetaxel as second-line treatment for patients with advanced non-small-cell lung cancer (ZODIAC): a double-blind, randomised, phase 3 trial. Lancet Oncol. 2010;11(7):619–26.

2. Kaul U, Bangalore S, Seth A, Priyadarshini A, Rajpal KA, Tejas MP et al. Paclitaxel-Eluting versus EverolimusEluting Coronary Stents in Diabetes. N Engl J Med. 2015;373(18):1709-19.



```{r, echo = FALSE}
rm(list = ls())
```