---
title: "CompARE"
subtitle: "Welcome to CompARE - Designing feasible clinical trials"
output:
  html_document:
    theme: cosmo
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
h1.title {
  display: none;
}
</style>
```
```{=html}
<style type="text/css">
.subtitle {
  font-size: 36px;
  font-weight: 500;
  color: DarkRed;
  text-align: center;
}
</style>
```
```{=html}
<style type="text/css">
  body{
  font-size: 12pt;
}
</style>
```
```{=html}
<head>
<title>CompARE</title>
</head>
```
<br>

```{r echo=FALSE}
# https://stackoverflow.com/questions/61313078/create-a-carousel-in-rmarkdown
objpath <- file.path("images/carousel_home")
slickR::slickR(
    list.files(objpath,full.names = TRUE,pattern = 'png'),
    height = 300,
    width = '95%')
```

```{r echo=FALSE}
library(fontawesome)
```

<br>

<br>

### What is CompARE?

<hr>

**CompARE** is an initiative inspired to provide help on issues relating
to trials with composite endpoints. **CompARE** may be used as a tool
for calculating the elements needed in the planning phase of clinical
trials involving composite endpoints.

<br>

### Shiny Apps

<hr>

**CompARE** is split into two apps for time-to-event and binary
endpoints, respectively. They are implemented with the Shiny R package.
With its user-friendly interface, CompARE allows to input the main
parameters included in the trial -such as the treatment effect on the
components of the composite endpoint, and its frequencies of occurrence-
and helps provide power and sample size calculations among others. <br>

<ul class="pager">

<!--this is the style of the button-->

<li><a href="https://www.grbio.eu/compare/CompARETimeToEvent/">      TIME-TO-EVENT
APP      </a></li>

<li><a href="https://www.grbio.eu/compare/CompAREBinary/">      BINARY
APP      </a></li>

</ul>

<ul class="pager">

<!--this is the style of the button-->

<li><a href="https://www.grbio.eu/compare/CompARETimeToEvent/help_Tutorial.html">TIME-TO-EVENT
TUTORIAL</a></li>

<li><a href="https://www.grbio.eu/compare/CompAREBinary/Help-Tutorial.html">BINARY
TUTORIAL</a></li>

</ul>

<br>

### R Package - CompAREdesign

<hr>

You can also find the R-package version of CompARE in CRAN
[CompAREdesign](https://cran.r-project.org/web/packages/CompAREdesign/index.html).

It has been designed to calculate the required sample size in randomized
clinical trials with composite endpoints. This package also includes
functions to calculate the probability of observing the composite
endpoint and the expected effect on the composite endpoint, among
others.

```{=html}
<div id="visit-counter"></div>
```

```{=html}
<script>
  // Function to fetch visit count from Google Analytics
  function fetchVisitCount() {
    gtag('event', 'page_view', {
      'send_to': 'G-EGNE926NLF', // Replace with your tracking ID
      'event_callback': function() {
        var visitCount = gtag.getAll()[0].get('metric1'); // Adjust the metric number as per your configuration
        document.getElementById('visit-counter').innerText = 'Total Visits: ' + visitCount;
      }
    });
  }

  // Call the function to fetch visit count when the page loads
  window.onload = fetchVisitCount;
</script>
```
