--- 
title: "CompARE" 
subtitle: "Welcome to CompARE - Designing feasible clinical trials" 
output: 
  html_document: 
    theme: cosmo 
editor_options:  
  markdown:  
    wrap: 72 
--- 
 
```{=html} 
<style type="text/css"> 
h1.title { 
  display: none; 
} 
</style> 
``` 
```{=html} 
<style type="text/css"> 
.subtitle { 
  font-size: 36px; 
  font-weight: 500; 
  color: DarkRed; 
  text-align: center; 
} 
</style> 
``` 
```{=html} 
<style type="text/css"> 
  body{ 
  font-size: 12pt; 
} 
</style> 
``` 
```{=html} 
<head> 
<title>CompARE</title> 
</head> 
``` 
<br> 
 
```{r echo=FALSE} 
# https://stackoverflow.com/questions/61313078/create-a-carousel-in-rmarkdown 
objpath <- file.path("images/carousel_home") 
slickR::slickR( 
    list.files(objpath,full.names = TRUE,pattern = 'png'), 
    height = 300, 
    width = '95%') 
``` 
 
```{r echo=FALSE} 
library(fontawesome) 
``` 
 
<br> 
 

```{r Calling-URL, include=FALSE}

library(gsheet)

# URL
url <- "https://docs.google.com/spreadsheets/d/1-Vklh823msjU9esHyUXF9v6uB8YXULZ_CAvB1bspjhA/edit?usp=sharing"



DF <- gsheet2tbl(url)
totalusers <- as.numeric(DF[11,1])


################################

# Fet amb lel compte de GRBIO
# Shiny Apps

url2 <- "https://docs.google.com/spreadsheets/d/1NPnym2QZvmyivwP8R8u2y00KYHQJhIi0u-m9eOIRkxw/edit?usp=sharing"

DF2 <- gsheet2tbl(url2)
totalusers_tte <- as.numeric(DF2[17, 2])+ as.numeric(DF2[18, 2])
totalusers_bin <- as.numeric(DF2[16, 2])

# Web


```


<br> 
 
### What is CompARE? 
 
<hr> 
 
**CompARE** is an initiative inspired to provide help on issues relating 
to trials with composite endpoints. **CompARE** may be used as a tool 
for calculating the elements needed in the planning phase of clinical 
trials involving composite endpoints. 
 
<br> 
 
### Shiny Apps 
 
<hr> 
 
**CompARE** is split into two apps for time-to-event and binary 
endpoints, respectively. They are implemented with the Shiny R package. 
With its user-friendly interface, CompARE allows to input the main 
parameters included in the trial -such as the treatment effect on the 
components of the composite endpoint, and its frequencies of occurrence- 
and helps provide power and sample size calculations among others. <br> 
 
<ul class="pager"> 
 
<!--this is the style of the button--> 
 
<li><a href="https://www.grbio.eu/compare/CompARETimeToEvent/">      TIME-TO-EVENT 
APP      </a></li> 
 
<li><a href="https://www.grbio.eu/compare/CompAREBinary/">      BINARY 
APP      </a></li> 
 
</ul> 
 
<ul class="pager"> 
 
<!--this is the style of the button--> 
 
<li><a href="https://www.grbio.eu/compare/CompARETimeToEvent/help_Tutorial.html">TIME-TO-EVENT 
TUTORIAL</a></li> 
 
<li><a href="https://www.grbio.eu/compare/CompAREBinary/Help-Tutorial.html">BINARY 
TUTORIAL</a></li> 
 
</ul> 
 
<br> 

![](https://img.shields.io/badge/Total Users Binary App-`r totalusers_tte`-blue)
![](https://img.shields.io/badge/Total Users Time to Event App-`r totalusers_bin`-blue)

<br>
 
### R Package - CompAREdesign 
 
<hr> 
 
You can also find the R-package version of CompARE in CRAN 
[CompAREdesign](https://cran.r-project.org/web/packages/CompAREdesign/index.html). 
 
It has been designed to calculate the required sample size in randomized 
clinical trials with composite endpoints. This package also includes 
functions to calculate the probability of observing the composite 
endpoint and the expected effect on the composite endpoint, among 
others. 
 
<br> 
 


          
![](https://img.shields.io/badge/Total Users Website-`r totalusers`-blue)



 
