---
title: "Case studies with CompAREdesign: Time-to-event endpoint"
author: "Marta Bofill Roig, Jordi Cortés Martínez and Guadalupe Gómez Melis"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE, echo=FALSE}
library('CompAREdesign')            # Load library
knitr::opts_chunk$set(echo = TRUE)  # Show code
```


## Time-to-event endpoint: ZODIAC TRIAL

This example is based on the data from the [ZODIAC trial](https://doi.org/10.1016/S1470-2045(10)70132-7) [1].

### Input Parameters

First of all, the information for the components of the composite endpoint should be defined.

```{r}
## Probabilities of observing the event in control arm during follow-up
p0_e1 <- 0.59   # Death
p0_e2 <- 0.74   # Disease Progression  

## Effect size (Cause specific hazard ratios) for each endpoint
HR_e1 <- 0.91   # Death
HR_e2 <- 0.77   # Disease Progression

## Hazard rates over time
beta_e1 <- 2    # Death --> Increasing risk over time
beta_e2 <- 1    # Disease Progression --> Constant risk over time

## Correlation
rho      <- 0.1        # Correlation between components
rho_type <- 'Spearman' # Type of correlation measure
copula   <- 'Frank'    # Copula used to get the joint distribution

## Additional parameter
case  <- 3  # 1: No deaths;                  2: Death is the secondary event; 
            # 3: Death is the primary event; 4: Both events are death by different causes
```

### ARE: Asymptotic Relatively Efficiency

As the ARE is greater than 1, the design using the composite endpoint as the primary endpoint is more efficient from the statistical point of view.

```{r}
ARE_tte(p0_e1   = p0_e1    , p0_e2    = p0_e2, 
        HR_e1   = HR_e1    , HR_e2    = HR_e2, 
        beta_e1 = beta_e1  , beta_e2  = beta_e2,  
        rho     = rho      , rho_type = rho_type, 
        copula  = copula   , case     = case) 
```

### Effect size of the Composite Endpoint

Several summary measures of the treatment effect are provided.

```{r}
effectsize_tte(p0_e1   = p0_e1    , p0_e2    = p0_e2, 
               HR_e1   = HR_e1    , HR_e2    = HR_e2, 
               beta_e1 = beta_e1  , beta_e2  = beta_e2,  
               rho     = rho      , rho_type = rho_type, 
               copula  = copula   , case     = case) 
```

### Sample size

The required number of patients for the design of the trial using the composite endpoint as the primary endpoint is **1118**.

```{r}
samplesize_tte(p0_e1   = p0_e1    , p0_e2    = p0_e2, 
               HR_e1   = HR_e1    , HR_e2    = HR_e2, 
               beta_e1 = beta_e1  , beta_e2  = beta_e2,  
               rho     = rho      , rho_type = rho_type, 
               copula  = copula   , case     = case, 
               alpha   = 0.025    , power    = 0.90,
               ss_formula = 'schoenfeld') 

```

### Influence of hazards rates over time on the effect size

The influence of the behaviour of the hazard rates over time on the treatment effect
can be studied by the function `effectsize_tte`.

```{r}
## Hazard rates over time Case 1
beta_e1 <- 1    # Death --> constant over time
beta_e2 <- 2    # Disease Progression --> increase over time
effectsize_tte(p0_e1   = p0_e1    , p0_e2    = p0_e2, 
               HR_e1   = HR_e1    , HR_e2    = HR_e2, 
               beta_e1 = beta_e1  , beta_e2  = beta_e2,  
               rho     = rho      , rho_type = rho_type, 
               copula  = copula   , case     = case) 

## Hazard rates over time Case 2
beta_e1 <- 1    # Death --> constant over time
beta_e2 <- 1    # Disease Progression --> constant over time
effectsize_tte(p0_e1   = p0_e1    , p0_e2    = p0_e2, 
               HR_e1   = HR_e1    , HR_e2    = HR_e2, 
               beta_e1 = beta_e1  , beta_e2  = beta_e2,  
               rho     = rho      , rho_type = rho_type, 
               copula  = copula   , case     = case) 
```


## References

1. Herbst RS, Sun Y, Eberhardt WEE, Germonpré P, Saijo N, Zhou C et al. Vandetanib plus docetaxel versus docetaxel as second-line treatment for patients with advanced non-small-cell lung cancer (ZODIAC): a double-blind, randomised, phase 3 trial. Lancet Oncol. 2010;11(7):619–26.

2. Kaul U, Bangalore S, Seth A, Priyadarshini A, Rajpal KA, Tejas MP et al. Paclitaxel-Eluting versus EverolimusEluting Coronary Stents in Diabetes. N Engl J Med. 2015;373(18):1709-19.



```{r, echo = FALSE}
rm(list = ls())
```